---
output: github_document
---

# using R to write a MATLAB function file

I want to create MATLAB function files from R to support the reproducibility of MATLAB course materials written in R Markdown. 

The idea is to: 

- write the lines of a MATLAB user-defined function as a R string 
- in R, write those lines to a text file `function.m` 
- eventually write a single R script that creates all my commonly used m-functions that I can `source()` from any Rmd 

In MATLAB R2016b or later, multiple functions can be defined in a single script. However, my current plan is to create separate m-files for each function. 



## managing files 

I use relative file paths with respect to the RStudio Project working directory. My examples require a file structure that includes:  

```
project\
  |-- derived\
  |-- results\
  |-- scripts\
  `-- project.Rproj
```

If you want a complete script to execute right now, copy and save [02_functions_script.Rmd](02_functions_script.Rmd) as an Rmd script in the `scripts\` directory and knit. 



## getting started 

Create an Rmd script and save it to the `scripts/` directory. 

```{r echo = 1:2}
library(knitr)
opts_knit$set(root.dir = "../")
opts_chunk$set(echo = TRUE, comment = NA, message = FALSE, warning = FALSE, collapse = FALSE, fig.keep = 'high')
```

Packages. 

```{r echo = TRUE}
library(readr)
library(reach)
library(R.matlab)
```

Executed only if a MATLAB path error occurs. 

```{r echo = TRUE, eval = FALSE}
add_to_path <- "pathstr = [cd]; 
  addpath(genpath(pathstr), '-end'); 
  savepath;" 
reach::runMatlabCommand(add_to_path)
```



## write_sys()

My test-case MATLAB function, `write_sys.m`, writes a system transfer function to a text file. The function is based on the following code chunk from the Bode plot tutorial:    

```
% write sys to txt
fid = fopen('results/sys_tf.txt', 'w');
tfString = evalc('sys');
fprintf(fid, '%s', tfString);
fclose(fid);
```

I formulate formulate these lines as a MATLAB function in a string in the Rmd script.  Use `cat()` to write the lines to a text file with a `.m` suffix in the `derived/` directory. 

```{r}
# user-defined MATLAB function
function_lines <- "function write_sys(sys, path)
  fid = fopen(path, 'wt');
  tf_string = evalc('sys');
  fprintf(fid, tf_string);
  fclose(fid);
  end
"
# write to file 
cat(function_lines, file = 'derived/write_sys.m', sep = '\n', append = FALSE)
```

The arguments are: 

- `sys` the result of the MATLAB `tf()` function 
- `path` the relative path and filename to be written 

The `scripts/` directory should include this Rmd script. The `derived/` directory should have the m-file we just made.



## test the function 

I'll create a transfer function for a first-order system and include the new  function call, writing the transfer function to the `results/` directory. 

```{r}
m_script <- "% assign parameters
K  = 1;
wb = 0.5;

% create the transfer function 
n = K;
d = [1/wb  1];
sys = tf(n, d);

% write sys to txt
write_sys(sys, 'results/sys02.txt')
"
```

Then execute the m-file. 

```{r eval = FALSE}
# run the m-file
reach::runMatlabCommand(m_script)
Sys.sleep(12)
```

Examine the results. 

```{r}
# read the lines saved by the function 
sys <- read_lines('results/sys02.txt', skip = 3, n_max = 3)

# print to the document
cat(sys, sep = "\n")
```



## session info

```{r}
library(devtools)
session_info()
```

--- 
[main page](../README.md)
