---
output: github_document
---

# using R to write a MATLAB function file

I want to create MATLAB function files from R: 

- In an R string, write the lines of a MATLAB user-defined function 
- In R, write those lines to a text file `function.m` 
- The function is then accessible to MATLAB scripts executed from R 
- Functions I use regularly can be collected in a single R script that I `source()` from any Rmd. 
- While functions are supported in scripts in MATLAB R2016b or later, my current plan is to write each function to separate m-file. 

My goal is to support the reproducibility of MATLAB course materials written in R Markdown. 

## getting started 

Create an Rmd script and save it to the `scripts/` directory. 

```{r echo = 1:2}
library(knitr)
opts_knit$set(root.dir = "../")
opts_chunk$set(echo = TRUE, comment = NA, message = FALSE, warning = FALSE, collapse = FALSE, fig.keep = 'high')
```

Packages. 

```{r echo = TRUE}
library(readr)
library(reach)
library(R.matlab)
```

Executed only if a MATLAB path error occurs. 

```{r echo = TRUE, eval = FALSE}
add_to_path <- "pathstr = [cd]; 
  addpath(genpath(pathstr), '-end'); 
  savepath;" 
reach::runMatlabCommand(add_to_path)
```


## write_sys()

My test-case MATLAB function writes a system transfer function to a text file. The function is based on code from the Bode plot tutorial:    

```
% write sys to txt
fid = fopen('results/sys_tf.txt', 'w');
tfString = evalc('sys');
fprintf(fid, '%s', tfString);
fclose(fid);
```

Formulating these lines as a MATLAB function in a string in the Rmd script and writing the lines to a text file with a `.m` suffix.  

```{r}
# used-defined MATLAB function
function_lines <- "function write_sys(sys, path)
  fid = fopen(path, 'wt');
  tf_string = evalc('sys');
  fprintf(fid, tf_string);
  fclose(fid);
end
"

# write to file 
cat(function_lines, file = 'scripts/write_sys.m', sep = '\n', append = FALSE)
```

The arguments are: 

- `sys` the result of the MATLAB `tf()` function 
- `path` the relative path and filename to be written  

The `scripts/` directory should include this Rmd script and the m-file we just made.

```
scripts\
  |-- your_filename.Rmd
  `-- write_sys.m 
```

## test the function 

I'll create a transfer function for a first-order system and include the new  function call: 

```{r}
m_script <- "% assign parameters
K  = 1;
wb = 0.5;

% create the transfer function 
n = K;
d = [1/wb  1];
sys = tf(n, d);

% write sys to txt
write_sys(sys, 'results/sys02.txt')
"
```

Then execute the m-file: 

```{r eval = FALSE}
reach::runMatlabCommand(m_script)
Sys.sleep(12)
```

Read the text file produced by the function and print it. 

```{r}
sys <- read_lines('results/sys02.txt', skip = 3, n_max = 3)
cat(sys, sep = "\n")
```

## session info

```{r}
library(devtools)
session_info()
```

--- 
[main page](../README.md)
